\chapter{Budget} 
\label{chap:economic}

This project has been carried out and funded with the Ericsson-UPM scholarship in software and systems.
from July 2020 to Jun 2021. The budget for this project are labor costs, computer equipment
and software costs.

\begin{itemize}
  \item \textbf{Salaries}
  
  This project was mainly developed by the student and the supervisor. One to three one-hour sessions a week is 
  done to see the progress, so the deditation of the supervisor is calculated based on that.
  For reference, the average salary of a software engineer in Stockholm \cite{glsd}, Swiden is 44000 SEK/month which 
  is equivalent to 4346,43 EUR at the time of writing. 

  The student is paid 500 EUR/month with a 20 hours/week dedication during 12 months.

  \item \textbf{Equipment and Software}
  
  The system used to develope this project is provided by Ericsson, it is a HP EliteBook 850 G6 with a cost of 1999 EUR
  , and it has this specifications:
  \begin{itemize}
    \item [$\circ$] Intel Core i7-8665U CPU @ 1.90GHz
    \item [$\circ$] 32GB RAM
    \item [$\circ$] 512GB PCIe NVMe M.2 SSD
    \item [$\circ$] Windows 10 Enterprice
  \end{itemize}

  The paid software include Microsoft Office Suite, with a cost of 149 EUR.
\end{itemize}

\begin{table}[]
  \begin{tabular}{@{}lllll@{}}
  \toprule
             & Salary/month (EUR) & \multicolumn{2}{l}{Months}                                                     & Cost (EUR)  \\ \midrule
  Student    & 500                & \multicolumn{2}{l}{12}                                                         & 6000  \\ \bottomrule
             & Salary (EUR)       & \begin{tabular}[c]{@{}l@{}}Salary per hour\\ (EUR)\end{tabular} & Worked hours & Cost (EUR)   \\ \midrule
  Supervisor & 4346.43            & 32.6                                                              & 100          & 3260  \\ \bottomrule
             & Cost               & \multicolumn{2}{l}{Amortization years}                                         & Cost (EUR)   \\ \midrule
  Laptop     & 1999               & \multicolumn{2}{l}{10}                                                         & 199   \\
  MS Office  & 149                & \multicolumn{2}{l}{1}                                                          & 149   \\ \bottomrule
   & & & \textbf{TOTAL}                                                                                       & \textbf{9608 EUR} \\ \bottomrule
  \end{tabular}
  \caption{Budget}
  \end{table}

  The Total cost of this project is 9608 EUR in total.


